---
title: "6. Supplementary Figure - Total ARGs"
author: "Avril Metcalfe-Roach"
date: "`r Sys.Date()`"
output: html_document
---

```{r Load Data}
library(tidyverse)
library(ggpubr)

# Load ARG master file from CARD database
card = readxl::read_xlsx('Data Files/CARD/P-00IK_Antibiotic/ARG_master.xlsx') %>% 
  rename(MDB_ID = SampleID) %>% 
  select(c(Protein_Accession:Resistance_Mechanism))

# Load phyloseq object & extract sample info
ps = readRDS('Data Files/bacterial_phyloseq.rds')
psm = ps %>% psmelt() %>% select(-Abundance, -OTU, -c(Kingdom:Species)) %>% 
  unique %>% left_join(card) %>% 
  mutate(Resistance_Mechanism = str_to_title(Resistance_Mechanism) %>% str_wrap(30))

# Calculate total ARGs
df = psm %>% filter(Day == 'Day 28') %>% 
  group_by(Sample,Diet) %>% 
  summarize(Count = sum(Count)) %>% ungroup()
# For Day 0 as well
df.all = psm %>%
  group_by(Sample,Diet,Day) %>% 
  summarize(Count = sum(Count)) %>% ungroup()

phylum = ps %>% tax_glom('Phylum') %>% microbiome::transform('compositional') %>% psmelt() %>% 
  select(Sample,Phylum,Abundance) %>% rename(phy_ab = Abundance)
phylum.mean = phylum %>% 
  group_by(Phylum) %>% summarize(phy_ab = mean(phy_ab)) %>% ungroup %>% 
  filter(phy_ab >0.01)
phylum = phylum %>% filter(Phylum %in% phylum.mean$Phylum)
```

```{r Prep Pathobionts}
# Pathobionts of interest
# Genus level:
path.genus = c('Streptococcus','Staphylococcus','Enterococcus','Escherichia','Shigella','Haemophilus','Campylobacter','Klebsiella')
# Species level:
# NOTE: phocaeicola vulgatus = bacteroides vulgatus
path.species = c('Phocaeicola vulgatus', 'Prevotella copri','Bacteroides fragilis')
  
# Load using a function so that ps isn't overwritten
load_path = function(){
  ps = readRDS('Data Files/bacterial_phyloseq.rds')
  ps.rel = ps %>% microbiome::transform('compositional')
  
  # Maaslin stats
  maas = readxl::read_xlsx('Results Files/Bacteriome/Maaslin2_longitudinal_bacteria.xlsx',sheet='All Results')

  # Extract otu data into table
  g = ps.rel %>% subset_taxa(Genus %in% path.genus) %>% tax_glom('Genus') %>% 
    psmelt() %>% rename(Pathobiont = Genus) %>% select(-OTU) %>% select(-c(Kingdom:Family))
  sp = ps.rel %>% subset_taxa(Species %in% path.species) %>% tax_glom('Species') %>% 
    psmelt() %>% rename(Pathobiont = Species) %>% select(-OTU) %>% select(-c(Kingdom:Genus))
  
  maas2 = maas %>% filter((Level=='Genus' & Genus %in% path.genus) | (Level=='Species' & Species %in% path.species)) %>% 
    mutate(Pathobiont = ifelse(is.na(Species),Genus,Species)) %>% 
    select(Pathobiont, qval) %>% 
    mutate(symb = ifelse(qval>0.05,'',ifelse(qval>0.01,"  * ",
                        ifelse(qval>0.001,' ** ',ifelse(qval>0.0001,' ***','****')))))
  
  # sample-specific data with qvals
  path=rbind(g,sp) %>% left_join(maas2)
  return(path)
}

path = load_path()

# Categorize pathobionts
path$cat = sapply(path$Pathobiont,function(x)ifelse(x %in% c('Staphylococcus','Streptococcus','Prevotella copri'),'Down','Up'))
```

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Total Pathobionts - all groups

```{r}
my_comparisons <- list( c("CON Day 0", "LM Day 0"), 
                        c("CON Day 0", "CON Day 28"),
                        c("LM Day 0", "LM Day 28"), 
                        c("LM Day 28", "CON Day 28") )

path %>% 
  group_by(Sample,Day,Diet,Group) %>% 
  summarise(Abundance = sum(Abundance)) %>% ungroup() %>% 
  ggplot(aes(Group, Abundance, fill = Group)) +
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(height = 0, width = 0.2) +
  theme_classic(base_size = 16) + 
  ggpubr::stat_compare_means(comparisons = my_comparisons,size=5)+
  ylab('Total Pathobionts (Rel Ab)') +
  xlab('') + theme(legend.position = 'none')
ggsave('Plots/Bacteriome/S1_totalpath_boxplots.jpeg',height = 5, width = 6)
```

# Total ARG vs Total Pathobionts

```{r}
annotations = path %>% filter(Day=='Day 28') %>% 
  group_by(Sample,Diet,Group) %>% 
  summarise(Abundance = sum(Abundance)) %>% ungroup() %>% 
  left_join(df) %>% 
  group_by(Diet) %>% 
  group_modify(~cor.test(.$Abundance,.$Count,method='spearman')[c("statistic","p.value","estimate")] %>% as.data.frame()) %>% 
  mutate(label.lm = paste(Diet,': \u03C1=',as.character(round(estimate,2)),
                          ', p=',signif(p.value,2),sep='')) %>%
  mutate(label.con = ifelse(Diet=='CON',label.lm,NA),
         label.lm = ifelse(Diet=='LM',label.lm,NA))

path %>% filter(Day=='Day 28') %>% 
  group_by(Sample,Diet,Group) %>% 
  summarise(Abundance = sum(Abundance)) %>% ungroup() %>% 
  left_join(df) %>% 
  ggplot(aes(Abundance,Count, fill = Diet,col=Diet)) +
  geom_point() +
  geom_smooth(method='lm') +
  theme_classic(base_size = 16) + 
  geom_text(aes(x=0,y=750,label=label.lm),
            data = annotations, size = 4.5,lineheight=0.8,hjust=0) +
  geom_text(aes(x=0,y=680,label=label.con),
            data = annotations, size = 4.5,lineheight=0.8,hjust=0) +
  xlab('Total Pathobionts (Rel Ab)') +
  ylab('Total ARGs (RPKM)') + theme(legend.position = 'none')
ggsave('Plots/Bacteriome/S1_totalpath_totalARG.jpeg',height = 5, width = 5)
```

# Total ARGs vs individual pathobionts

```{r}
temp = path %>% filter(Day=='Day 28') %>% 
  filter(symb != '') %>% # Only significant taxa
  left_join(df)

# CON
df.stats.con = temp %>% filter(Diet=='CON') %>% 
  group_by(Pathobiont,cat) %>% 
  group_modify(~cor.test(.$Abundance,.$Count,method='spearman')[c("statistic","p.value","estimate")] %>% as.data.frame()) %>% ungroup() %>% 
  mutate(qval = p.adjust(p.value,method='BH')) %>% ungroup %>% 
  mutate(Group = 'CON')
# LM
df.stats.lm = temp %>% filter(Diet=='LM') %>% 
  group_by(Pathobiont,cat) %>% 
  group_modify(~cor.test(.$Abundance,.$Count,method='spearman')[c("statistic","p.value","estimate")] %>% as.data.frame()) %>% ungroup() %>% 
  mutate(qval = p.adjust(p.value,method='BH')) %>% ungroup %>% 
  mutate(Group = 'LM')

df.stats = rbind(df.stats.con,df.stats.lm) %>% filter(!is.na(estimate))
write.csv(df.stats,'Results Files/Functional/Indiv_pathobiont_vs_total_ARG.csv',row.names = F)
```

# Total ARGs vs Phyla

```{r}
temp = df.all %>% left_join(phylum)

# CON
df.stats.con = temp %>% filter(Diet=='CON') %>% 
  group_by(Phylum,Day) %>% 
  group_modify(~cor.test(.$phy_ab,.$Count,method='spearman')[c("statistic","p.value","estimate")] %>% 
                 as.data.frame()) %>% ungroup() %>% 
  group_by(Day) %>% 
  mutate(qval = p.adjust(p.value,method='BH')) %>% ungroup %>% 
  mutate(Group = 'CON')
# LM
df.stats.lm = temp %>% filter(Diet=='LM') %>% 
  group_by(Phylum,Day) %>% 
  group_modify(~cor.test(.$phy_ab,.$Count,method='spearman')[c("statistic","p.value","estimate")] %>% 
                 as.data.frame()) %>% ungroup() %>% 
  group_by(Day) %>% 
  mutate(qval = p.adjust(p.value,method='BH')) %>% ungroup %>% 
  mutate(Group = 'LM')

df.stats = rbind(df.stats.con,df.stats.lm) %>% filter(!is.na(estimate))
write.csv(df.stats,'Results Files/Functional/Phyla_vs_total_ARG.csv',row.names = F)

# ~~~~~~ Plots

annotations = df.stats %>% 
  mutate(label.lm = paste(Group,': \u03C1=',as.character(round(estimate,2)),', q=',signif(qval,2),sep='')) %>%
  mutate(label.con = ifelse(Group=='CON',label.lm,NA),
         label.lm = ifelse(Group=='LM',label.lm,NA)) %>% 
  select(Phylum,Day,Group,label.lm,label.con) %>% rename(Diet=Group) %>% 
  mutate(ylab = ifelse(Day=='Day 0' & Diet=='CON',250,
                       ifelse(Day=='Day 0' & Diet=='LM',280,
                              ifelse(Day=='Day 28' & Diet=='CON',700,
                                  ifelse(Day=='Day 28' & Diet=='LM',780,NA)))))

temp %>% left_join(annotations) %>% 
  ggplot(aes(phy_ab,Count, fill = Diet,col=Diet)) +
  geom_point() +
  geom_smooth(method='lm') +
  geom_text(aes(x=0.01,y=ylab,label=label.lm,fill=Diet,col=Diet),
            data = annotations, hjust=0,size=3.5) +
  geom_text(aes(x=0.01,y=ylab,label=label.con,fill=Diet,col=Diet),
            data = annotations, hjust=0,size=3.5) +
  theme_classic(base_size = 16) + 
  ggh4x::facet_nested(Day ~ Phylum,
                      scales = 'free',space='fixed',axes='all') +
  xlab('Phylum Relative Abundance') +
  ylab('Total ARGs (RPKM)') + theme(legend.position = 'none')

ggsave('Plots/Bacteriome/S1_phylum_totalARG.jpeg',height = 5, width = 10)
```

