---
title: "6. Figure 4 with CARD Database"
author: "Avril Metcalfe-Roach"
date: "`r Sys.Date()`"
output: html_document
---

```{r Load Data}
library(phyloseq)
library(tidyverse)
library(ggpubr)
```

```{r Load SEED Data}
# Load data
df = read.csv("Data Files/SEED_abundances.csv")

# Taxonomy
t = df %>% select(contains('Subsystem'),Function) %>% 
  mutate(OTU = paste(Subsystem.Level.1,Subsystem.Level.2,Subsystem.Level.3, Function,sep='|')) %>% 
  mutate_all(function(x)str_replace_all(x,"'",'.')) %>% 
  column_to_rownames('OTU') %>% as.matrix %>% tax_table()

# Abundance
o = df %>% select(-all_of(colnames(t))) %>% as.matrix %>% `rownames<-`(rownames(t))

# Convert to TPM
cs = colSums(o)/1e6 # Denominator
o2 = (t(o)/cs) %>% t()

# Sample data
s = read.csv('Data Files/metadata.csv') %>% `rownames<-`(.$Sample) %>% 
  mutate(Mouse = str_remove(Sample,'_D0') %>% str_remove('_D28'))

# Phyloseq object
ps = phyloseq(sample_data(s),otu_table(o2,taxa_are_rows=T),t)
ps@sam_data$Group = factor(ps@sam_data$Group, levels = c('CON Day 0','LM Day 0','CON Day 28','LM Day 28'))

cols = c('CON' = '#FC746E','LM' = '#01BEC3')

# SEED Maaslin2 results
maas = readxl::read_xlsx('Results Files/Functional/Maaslin2_SEED_D28.xlsx',sheet = 'All Results')
```

```{r Prep SEED Data}
temp1 = ps %>% subset_taxa(Function=='Integron_integrase_IntI1') %>% 
  psmelt() %>% 
  filter(Day=='Day 28') %>% mutate(label = 'IntI1') %>% 
  select(Sample,Day,Diet,Group,Abundance,label)
temp2 = ps %>% subset_taxa(Function=='Integron_integrase_IntI2') %>% 
  psmelt() %>% 
  filter(Day=='Day 28') %>% mutate(label = 'IntI2') %>% 
  select(Sample,Day,Diet,Group,Abundance,label)
temp3 = ps %>% subset_taxa(Subsystem.Level.3 == 'Conjugative transposon, Bacteroidales') %>%   
  tax_glom('Subsystem.Level.3') %>% 
  psmelt() %>% 
  filter(Day=='Day 28') %>% mutate(label = 'Conj Transp') %>% 
  select(Sample,Day,Diet,Group,Abundance,label)

temp = rbind(temp1,temp2,temp3)

lm_order = temp %>% filter(Diet=='LM') %>% 
  group_by(label) %>% 
  summarize(Abundance = mean(Abundance)) %>% ungroup() %>% 
  arrange(-Abundance)
```

```{r Prep CARD Data}
card = readxl::read_xlsx('Data Files/CARD/P-00IK_Antibiotic/ARG_master.xlsx') %>% 
  rename(MDB_ID = SampleID) %>% 
  select(c(Protein_Accession:Resistance_Mechanism))

# Add sample info
card = card %>% left_join(sample_data(ps) %>% as.data.frame)

# 30% prevalence filter
prevfilter = card %>% filter(Day == 'Day 28') %>% 
  group_by(Sample,ARO_Accession) %>% 
  summarize(Count = sum(Count)) %>% ungroup() %>% 
  mutate(Count = Count>0) %>% 
  group_by(ARO_Accession) %>% 
  summarize(Count = sum(Count)/20) %>% ungroup() %>% 
  filter(Count>0.3)
card = card %>% filter(ARO_Accession %in% prevfilter$ARO_Accession)
```

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Correlate ARGs with SEED

```{r Combine}
df = left_join(temp,card)
```

```{r Run stats and plot}
# CON
df.stats.con = df %>% filter(Diet=='CON',Day=='Day 28') %>% 
  group_by(label,ARO_Accession,Resistance_Mechanism) %>% 
  group_modify(~cor.test(.$Abundance,.$Count,method='spearman')[c("statistic","p.value","estimate")] %>% as.data.frame()) %>% 
  ungroup %>% group_by(label) %>% 
  mutate(qval = p.adjust(p.value,method='BH')) %>% ungroup %>% 
  mutate(Group = 'CON')
# LM
df.stats.lm = df %>% filter(Diet=='LM',Day=='Day 28') %>% 
  group_by(label,ARO_Accession,Resistance_Mechanism) %>% 
  group_modify(~cor.test(.$Abundance,.$Count,method='spearman')[c("statistic","p.value","estimate")] %>% as.data.frame()) %>% 
  ungroup %>% group_by(label) %>% 
  mutate(qval = p.adjust(p.value,method='BH')) %>% ungroup %>% 
  mutate(Group = 'LM')

df.stats = rbind(df.stats.con,df.stats.lm) %>% filter(!is.na(estimate)) %>% 
  mutate(label.lm = paste(Group,': \u03C1=',as.character(round(estimate,2)),', q=',signif(qval,2),sep='')) %>%
  mutate(label.con = ifelse(Group=='CON',label.lm,NA),
         label.lm = ifelse(Group=='LM',label.lm,NA))

# Arrange significance ~~~~~~~~~~~~~
df.stats %>% 
  mutate(ARO_Accession = as.character(ARO_Accession),
         Resistance_Mechanism = str_wrap(Resistance_Mechanism,25) %>% str_to_title) %>% 
  rename("Rho" = estimate) %>% 
  mutate(Qval = ifelse(qval>0.05,'',
                      ifelse(qval>0.01,'*',
                             ifelse(qval>0.001,'**',
                                    ifelse(qval<=0.001,'***',''))))) %>% 
  ggplot(aes(label,ARO_Accession,col = Rho)) +
  geom_point(size =14) +
  theme_classic(base_size = 14) +
  scale_color_gradient2(low="darkblue", high="darkred", guide="colorbar") +
  geom_text(aes(label=Qval),size = 9, col = 'white',nudge_y = -0.15) +
  ylab('ARG (ARO Accession)') + xlab(NULL) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_grid(cols = vars(Group),rows = vars(Resistance_Mechanism), scales = 'free',space = 'free') +
  theme(strip.text.y.right = element_text(angle = 0))
ggsave('Plots/Functional/SEED/P6_ARG_Heatmap_cnt_IntI.jpeg',height=12,width=7)
```

```{r Save stats}
write.csv(df.stats,'Results Files/Functional/ARG_vs_cnt_IntI_Spearman.csv',row.names = F)
```

