---
title: "9. Heatmap of Viruses vs ARGs"
author: "Avril Metcalfe-Roach"
date: "`r Sys.Date()`"
output: html_document
---

```{r Load Data}
library(phyloseq)
library(tidyverse)
library(ggpubr)
```

```{r Load viruses}
# ps output is bacterial, raw counts
source('0. Initialize Viral Data.R')
```

```{r}
# Maaslin stats
maas = readxl::read_xlsx('Results Files/Virome/Fig. 2 Supplement - Viral Maaslin2.xlsx',sheet='Genus') %>% filter(qval<0.01)

ps.rel = ps %>% tax_glom('Genus') %>% microbiome::transform('clr') %>% 
  subset_taxa(Genus %in% maas$Genus) %>% psmelt() %>% 
  filter(Day=='Day 28') %>% 
  left_join(maas %>% select(Genus,coef,qval)) %>% 
  mutate(symb = ifelse(qval>0.05,'',ifelse(qval>0.01,"  * ",
                      ifelse(qval>0.001,' ** ',ifelse(qval>0.0001,' ***','****')))))
```

```{r Prep CARD Data}
card = readxl::read_xlsx('Data Files/ARG_master.xlsx') %>% 
  rename(MDB_ID = SampleID) %>% 
  select(c(Protein_Accession:Resistance_Mechanism))

# Add sample info
card = card %>% left_join(sample_data(ps) %>% as.data.frame)

# 30% prevalence filter
prevfilter = card %>% filter(Day == 'Day 28') %>% 
  group_by(Sample,ARO_Accession) %>% 
  summarize(Count = sum(Count)) %>% ungroup() %>% 
  mutate(Count = Count>0) %>% 
  group_by(ARO_Accession) %>% 
  summarize(Count = sum(Count)/20) %>% ungroup() %>% 
  filter(Count>0.3)
card = card %>% filter(ARO_Accession %in% prevfilter$ARO_Accession)
```

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Correlate ARGs with SEED

```{r Combine}
df = left_join(ps.rel,card) %>% mutate(Association = ifelse(coef>0,'Up','Down'))
```

```{r Run stats and plot}
# CON
df.stats.con = df %>% filter(Diet=='CON',Day=='Day 28') %>% 
  group_by(Genus,ARO_Accession,Resistance_Mechanism,Association) %>% 
  group_modify(~cor.test(.$Abundance,.$Count,method='spearman')[c("statistic","p.value","estimate")] %>% as.data.frame()) %>% 
  ungroup %>% group_by(Genus,Association) %>% 
  mutate(qval = p.adjust(p.value,method='BH')) %>% ungroup %>% 
  mutate(Group = 'CON')
# LM
df.stats.lm = df %>% filter(Diet=='LM',Day=='Day 28') %>% 
  group_by(Genus,ARO_Accession,Resistance_Mechanism,Association) %>% 
  group_modify(~cor.test(.$Abundance,.$Count,method='spearman')[c("statistic","p.value","estimate")] %>% as.data.frame()) %>% 
  ungroup %>% group_by(Genus,Association) %>% 
  mutate(qval = p.adjust(p.value,method='BH')) %>% ungroup %>% 
  mutate(Group = 'LM')

df.stats = rbind(df.stats.con,df.stats.lm) %>% filter(!is.na(estimate)) %>% 
  mutate(label.lm = paste(Group,': \u03C1=',as.character(round(estimate,2)),', q=',signif(qval,2),sep='')) %>%
  mutate(label.con = ifelse(Group=='CON',label.lm,NA),
         label.lm = ifelse(Group=='LM',label.lm,NA)) %>% 
  arrange(estimate) %>% 
  mutate(ARO_Accession = as.character(ARO_Accession))

order = df.stats %>% filter(Group=='LM',Genus==df.stats$Genus[1]) %>% 
  arrange(-estimate)
df.stats = df.stats %>% mutate(ARO_Accession = factor(ARO_Accession, levels = order$ARO_Accession))

# Arrange significance ~~~~~~~~~~~~~
df.stats %>% 
  mutate(Resistance_Mechanism = str_wrap(Resistance_Mechanism,25) %>% str_to_title) %>% 
  rename("Rho" = estimate) %>% 
  mutate(Qval = ifelse(qval>0.05,'',
                      ifelse(qval>0.01,'*',
                             ifelse(qval>0.001,'**',
                                    ifelse(qval<=0.001,'***',''))))) %>% 
  ggplot(aes(Genus,ARO_Accession,col = Rho)) +
  geom_point(size =14) +
  theme_classic(base_size = 14) +
  scale_color_gradient2(low="darkblue", high="darkred", guide="colorbar") +
  geom_text(aes(label=Qval),size = 9, col = 'white',nudge_y = -0.15) +
  ylab('ARG (ARO Accession)') + xlab(NULL) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  ggh4x::facet_nested(Resistance_Mechanism ~ Group + Association, scales = 'free', space = 'free') +
  # facet_grid(cols = vars(Group),rows = vars(Resistance_Mechanism), scales = 'free',space = 'free') +
  theme(strip.text.y.right = element_text(angle = 0))
ggsave('Plots/Functional/Fig. S3d - ARG vs Viral Heatmap.jpeg',height=12,width=16)
```

```{r Save stats}
write.csv(df.stats,'Results Files/Functional/Fig. S3d ARG vs Viral Heatmap Statistics.csv',row.names = F)
```

