---
title: "4. Indicator Species"
author: "Avril Metcalfe-Roach"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(phyloseq)
library(tidyverse)
library(ggpubr)
library(indicspecies)
```

# BACTERIA ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

```{r}
# ps output is bacterial, raw counts
ps = readRDS('Data Files/bacterial_phyloseq.rds')

ps.rel = ps %>% microbiome::transform('compositional')

# Maaslin stats
maas = readxl::read_xlsx('Results Files/Bacteriome/Fig. 1 Supplement - Bacterial Maaslin2.xlsx',sheet='All Results')

# Melted and pivoted dataset
df = ps %>% subset_taxa(!is.na(Species)) %>%
  psmelt() %>% 
  select(-OTU, -c(Kingdom:Genus)) %>% 
  pivot_wider(names_from = Species, values_from = Abundance) %>% 
  # Combine Day 0s into one group
  mutate(grp = ifelse(Day == 'Day 0','Day 0',as.character(Group)))
```

```{r Indicator Analysis}
# Analysis groups:
grps = df$grp
# Create df of just abundances (raw counts)
df_ind = df %>% select(-c(Sample:IPITT.AUC),-grp) %>% 
  `rownames<-`(df$Sample)

# Choose only species above 0.001% mean abundance
df_rel = df_ind %>% apply(1,function(x)x/sum(x))
m = rowMeans(df_rel)
w = which(m>1e-5)
df_ind = df_ind %>% select(any_of(rownames(df_rel)[w]))
# Choose only species above 0.3 prevalence
df_prev = df_rel>0
m = rowSums(df_prev)/ncol(df_prev)
w = which(m>0.3)
df_ind = df_ind %>% select(any_of(rownames(df_rel)[w]))

# Run analysis
set.seed(711) # For reproducibility
indval = multipatt(df_ind, grps, control = how(nperm=999))
# All results, including combinations of groups. However, we want only the indicators that are unique to each group.
capture.output(summary(indval), file = 'Results Files/Bacteriome/Indicator Species Results Summary.txt')

# Isolate the Indicator species for each group.
# Select only high-quality indicators
ind = indval$sign %>% filter(p.value<0.05, stat>0.7) %>% 
  dplyr::select(-index) %>%
  rownames_to_column('Taxonomy_ID') %>%
  pivot_longer(cols = -c(Taxonomy_ID,stat,p.value), names_to = 'group', values_to = 'Ind' )
# Select indicators unique to a group
ind_pos = ind %>%
  filter(Ind == 1) %>% group_by(Taxonomy_ID) %>% 
  dplyr::count() %>% ungroup() %>%
  filter(n==1)
ind = ind %>% filter(Taxonomy_ID %in% ind_pos$Taxonomy_ID) %>%
  filter(Ind==1) %>% dplyr::select(-Ind)
# Remove prefix from groups
srv = Vectorize(str_remove)
ind$group = srv(ind$group, 's.')
# Number of Indicators in each group
table(ind$group)
ind = ind %>% arrange(group)

write.csv(ind, 'Results Files/Bacteriome/Indicator Species Statistics.csv',row.names=F)
```

# Fig. 3a - Indicator Species Dot Plot

```{r Indic Plot}
df_plot = df_rel %>% t() %>% as.data.frame() %>% rownames_to_column('Sample') %>% 
  left_join(df %>% select(Sample:Group) %>% unique) %>% 
  pivot_longer(cols = -c(Sample,sample_Sample,Diet,Day,Group),names_to = 'Taxonomy_ID', values_to = 'Abundance') %>% 
  left_join(ind %>% dplyr::rename(Indicator = group)) %>%
  filter(is.na(Indicator)==F) %>%
  dplyr::rename(Species = Taxonomy_ID) %>%
  arrange(Diet) %>% arrange(Day) %>% 
  mutate(indsp = ifelse(Day == 'Day 0','Day 0',as.character(Group)))
  
df_plot$indsp =  factor(df_plot$indsp, levels = c('Day 0','CON Day 28','LM Day 28'))

# Take mean values per group
df_ind_s_grp = df_plot %>% group_by(Indicator, indsp, Group,Species) %>% 
  summarize(Abundance = mean(Abundance)) %>% ungroup %>% 
  dplyr::rename('Mean Ab.' = Abundance) %>% 
  mutate(Indicator = factor(Indicator,levels = c('Day 0','CON Day 28','LM Day 28')))

p = df_ind_s_grp %>% 
  mutate(Species = str_wrap(Species,40)) %>% 
  arrange(Indicator) %>% 
  mutate(Species = factor(Species, levels = .$Species %>% unique())) %>% 
  ggplot(aes(x = Group, y = Species)) +
  geom_point(aes(color = Indicator, size = `Mean Ab.`), alpha = 0.5) +
  scale_color_manual(values = c("#E7B800", "#FC4E07", "#00AFBB")) +
  theme(text = element_text(size=16),
        panel.background = element_blank(),
        panel.grid.major = element_line(colour = "#bdbdbd", linetype = "dotted"),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 35, hjust = 1)) +
  scale_size(range = c(0.5, 12)) + xlab('')

ggsave('Plots/Bacteriome/Fig. 3a - Indicator Dot Plot.png', height = 5, width = 9)
```

# Fig. 3b-d - Taxa of Interest

```{r Plot taxa of interest}
comps <- list( c("CON Day 0", "LM Day 0"), 
                        c("CON Day 0", "CON Day 28"),
                        c("LM Day 0", "LM Day 28"), 
                        c("LM Day 28", "CON Day 28") )

psg = ps %>% tax_glom('Genus') %>% microbiome::transform('compositional') %>% 
  psmelt() %>% 
  filter(Genus %in% c('Corynebacterium','Klebsiella','Enterobacter'))

psg %>% filter(Genus=='Corynebacterium') %>% 
  ggplot(aes(Group, Abundance, fill = Group)) +
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(height = 0, width = 0.2) +
  theme_classic(base_size = 16) + 
  ggpubr::stat_compare_means(comparisons = comps,size=5)+
  ylab('Corynebacterium (Rel Ab.)') +
  xlab(NULL) +
  theme(legend.position = 'none')
ggsave('Plots/Bacteriome/Fig. 3b - Corynebacterium.png',height = 5, width = 6)

psg %>% filter(Genus=='Enterobacter') %>% 
  ggplot(aes(Group, Abundance, fill = Group)) +
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(height = 0, width = 0.2) +
  theme_classic(base_size = 16) + 
  ggpubr::stat_compare_means(comparisons = comps,size=5)+
  ylab('Enterobacter (Rel Ab.)') +
  xlab(NULL) +
  theme(legend.position = 'none')
ggsave('Plots/Bacteriome/Fig. 3c - Enterobacter.png',height = 5, width = 6)

psg %>% filter(Genus=='Klebsiella') %>% 
  ggplot(aes(Group, Abundance, fill = Group)) +
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(height = 0, width = 0.2) +
  theme_classic(base_size = 16) + 
  ggpubr::stat_compare_means(comparisons = comps,size=5)+
  ylab('Klebsiella (Rel Ab.)') +
  xlab(NULL) +
  theme(legend.position = 'none')
ggsave('Plots/Bacteriome/Fig. 3d - Klebsiella.png',height = 5, width = 6)
```

# FUNGI ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

```{r}
# ps output is bacterial, raw counts
source('0. Initialize Fungal Data.R')

ps.rel = ps %>% microbiome::transform('compositional')

# Maaslin stats
maas = readxl::read_xlsx('Results Files/Fungome/Fig. 2 Supplement - Fungal Maaslin2.xlsx',sheet='All Results')

# Melted and pivoted dataset
df = ps %>% tax_glom('Species') %>% psmelt() %>% 
  select(-OTU, -c(Kingdom:Genus)) %>% 
  pivot_wider(names_from = Species, values_from = Abundance) %>% 
  # Combine Day 0s into one group
  mutate(grp = ifelse(Day == 'Day 0','Day 0',as.character(Group)))
```

```{r Indicator Analysis}
# Analysis groups:
grps = df$grp
# Create df of just abundances (raw counts)
df_ind = df %>% select(-c(Sample:MDB_ID),-grp) %>% 
  `rownames<-`(df$Sample)

# Choose only species above 0.001% mean abundance
df_rel = df_ind %>% apply(1,function(x)x/sum(x))
m = rowMeans(df_rel)
w = which(m>1e-5)
df_ind = df_ind %>% select(any_of(rownames(df_rel)[w]))
# Choose only species above 0.3 prevalence
df_prev = df_rel>0
m = rowSums(df_prev)/ncol(df_prev)
w = which(m>0.3)
df_ind = df_ind %>% select(any_of(rownames(df_rel)[w]))

# Run analysis
set.seed(711) # For reproducibility
indval = multipatt(df_ind, grps, control = how(nperm=999))
# All results, including combinations of groups. However, we want only the indicators that are unique to each group.
capture.output(summary(indval), file = 'Results Files/Fungome/Indicator Species Results Summary.txt')

# Isolate the Indicator species for each group.
# Select only high-quality indicators
ind = indval$sign %>% filter(p.value<0.05, stat>0.7) %>% 
  dplyr::select(-index) %>%
  rownames_to_column('Taxonomy_ID') %>%
  pivot_longer(cols = -c(Taxonomy_ID,stat,p.value), names_to = 'group', values_to = 'Ind' )
# Select indicators unique to a group
ind_pos = ind %>%
  filter(Ind == 1) %>% group_by(Taxonomy_ID) %>% 
  dplyr::count() %>% ungroup() %>%
  filter(n==1)
ind = ind %>% filter(Taxonomy_ID %in% ind_pos$Taxonomy_ID) %>%
  filter(Ind==1) %>% dplyr::select(-Ind)
# Remove prefix from groups
srv = Vectorize(str_remove)
ind$group = srv(ind$group, 's.')
# Number of Indicators in each group
table(ind$group)
ind = ind %>% arrange(group)

write.csv(ind, 'Results Files/Fungome/Indicator Species Statistics.csv',row.names=F)
```

The following indicators are depleted in one of D28 LM or CON (not baseline).
```{r Taxa of Interest}
# Isolate the Indicator species for each group.
# Select only high-quality indicators
ind = indval$sign %>% filter(p.value==0.001, stat>0.7) %>% 
  dplyr::select(-index) %>%
  rownames_to_column('Taxonomy_ID') %>%
  pivot_longer(cols = -c(Taxonomy_ID,stat,p.value), names_to = 'group', values_to = 'Ind' )
# Select indicators unique to a group
ind_pos = ind %>%
  filter(Ind == 1) %>% group_by(Taxonomy_ID) %>% 
  dplyr::count() %>% ungroup() %>%
  filter(n==2)
ind = ind %>% filter(Taxonomy_ID %in% ind_pos$Taxonomy_ID) %>%
  filter(Ind==0) %>% dplyr::select(-Ind)
# Remove prefix from groups
srv = Vectorize(str_remove)
ind$group = srv(ind$group, 's.')
# Number of Indicators in each group
table(ind$group)
ind = ind %>% arrange(group)

write.csv(ind, 'Results Files/Fungome/Indicator Species Statistics - Depleted in 1 Group.csv',row.names=F)
```

```{r Plot taxa of interest}
psg = ps %>% tax_glom('Species') %>% microbiome::transform('compositional') %>% 
  psmelt() %>% 
  filter(Species %in% ind$Taxonomy_ID) %>% 
  left_join(ind %>% select(Taxonomy_ID,group) %>% unique %>% 
              rename(Species = Taxonomy_ID, Indicator = group)) %>% 
  filter(Day =='Day 28') %>% 
  left_join(maas %>% filter(Level=='Species') %>% select(Species,coef,qval)) %>% 
  mutate(symb = ifelse(qval>0.05,'',ifelse(qval>0.01,"* ",
                      ifelse(qval>0.001,'**',ifelse(qval>0.0001,'***','****'))))) %>% 
  mutate(Species = str_wrap(Species,25))

# LM-ASSOCIATED ~~~~~~~~~~~~~~~~~~~~~

# Set order for indicators
order = psg %>% filter(Diet=='LM',Indicator=='LM Day 28') %>% 
  group_by(Species) %>% summarize(Abundance = mean(Abundance)) %>% ungroup() %>% 
  arrange(-Abundance)

# Set x and y coords for the p values
max_per_path = psg %>% group_by(Species) %>% summarize(M = 2*max(Abundance)) %>% ungroup %>% 
  mutate(Species = factor(Species, levels = order$Species)) %>% 
  filter(!is.na(Species))
max_pos = max_per_path %>% 
  arrange(Species) %>% mutate(N = c(1:nrow(.))+0.25) %>% 
  mutate(M_same = max(.$M))

p1 = psg %>% filter(Indicator == 'LM Day 28') %>% droplevels() %>% 
  mutate(Species = factor(Species,levels = order$Species)) %>% 
  left_join(max_pos) %>% 
  mutate(Abundance = Abundance + min(.$Abundance[.$Abundance !=0])/2) %>% 
  ggplot(aes(Species, Abundance, fill = Group)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position=position_jitterdodge(jitter.height=0,jitter.width = 0.2)) +
  theme_classic(base_size = 16) +
  ylab('Relative Abundance') +
  xlab(NULL) +
  geom_text(aes(x=N,y=M_same,label=symb), size = 7,lineheight=0.8,hjust=2) +
  scale_y_log10(expand = expansion(mult = c(0.1, 0.1))) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

# CON-ASSOCIATED ~~~~~~~~~~~~~~~~~~~~~

# Set order for indicators
order = psg %>% filter(Diet=='CON',Indicator=='CON Day 28') %>% 
  group_by(Species) %>% summarize(Abundance = mean(Abundance)) %>% ungroup() %>% 
  arrange(-Abundance)

# Set x and y coords for the p values
max_per_path = psg %>% group_by(Species) %>% summarize(M = 2*max(Abundance)) %>% ungroup %>% 
  mutate(Species = factor(Species, levels = order$Species)) %>% 
  filter(!is.na(Species))
max_pos = max_per_path %>% 
  arrange(Species) %>% mutate(N = c(1:nrow(.))+0.25) %>% 
  mutate(M_same = max(.$M))

p2 = psg %>% filter(Indicator == 'CON Day 28') %>% droplevels() %>% 
  mutate(Species = factor(Species,levels = order$Species)) %>% 
  left_join(max_pos) %>% 
  mutate(Abundance = Abundance + min(.$Abundance[.$Abundance !=0])/2) %>% 
  ggplot(aes(Species, Abundance, fill = Group)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position=position_jitterdodge(jitter.height=0,jitter.width = 0.2)) +
  theme_classic(base_size = 16) +
  ylab('Relative Abundance') +
  xlab(NULL) +
  geom_text(aes(x=N,y=M_same,label=symb), size = 7,lineheight=0.8,hjust=1) +
  scale_y_log10(expand = expansion(mult = c(0.1, 0.1))) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

p = ggarrange(plotlist = list(p2,p1),common.legend = T,legend = 'right', widths = c(3,5))

ggsave('Plots/Fungome/Fig. 3e - Fungal Indicator Species.png',height = 6, width = 10.5)
```




# VIRUSES ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

```{r}
# ps output is bacterial, raw counts
source('0. Initialize Viral Data.R')

ps.rel = ps %>% microbiome::transform('compositional')

# Maaslin stats
maas = readxl::read_xlsx('Results Files/Virome/Fig. 2 Supplement - Viral Maaslin2.xlsx',sheet='All Results')

# Melted and pivoted dataset
df = ps %>% tax_glom('Species') %>% psmelt() %>% 
  select(-OTU, -c(Kingdom:Genus)) %>% 
  pivot_wider(names_from = Species, values_from = Abundance) %>% 
  # Combine Day 0s into one group
  mutate(grp = ifelse(Day == 'Day 0','Day 0',as.character(Group)))
```

```{r Indicator Analysis}
# Analysis groups:
grps = df$grp
# Create df of just abundances (raw counts)
df_ind = df %>% select(-c(Sample:MDB_ID),-grp) %>% 
  `rownames<-`(df$Sample)

# Choose only species above 0.001% mean abundance
df_rel = df_ind %>% apply(1,function(x)x/sum(x))
m = rowMeans(df_rel)
w = which(m>1e-5)
df_ind = df_ind %>% select(any_of(rownames(df_rel)[w]))
# Choose only species above 0.3 prevalence
df_prev = df_rel>0
m = rowSums(df_prev)/ncol(df_prev)
w = which(m>0.3)
df_ind = df_ind %>% select(any_of(rownames(df_rel)[w]))

# Run analysis
set.seed(711) # For reproducibility
indval = multipatt(df_ind, grps, control = how(nperm=999))
# All results, including combinations of groups. However, we want only the indicators that are unique to each group.
capture.output(summary(indval), file = 'Results Files/Virome/Indicator Species Results Summary.txt')

# Isolate the Indicator species for each group.
# Select only high-quality indicators
ind = indval$sign %>% filter(p.value<0.05, stat>0.7) %>% 
  dplyr::select(-index) %>%
  rownames_to_column('Taxonomy_ID') %>%
  pivot_longer(cols = -c(Taxonomy_ID,stat,p.value), names_to = 'group', values_to = 'Ind' )
# Select indicators unique to a group
ind_pos = ind %>%
  filter(Ind == 1) %>% group_by(Taxonomy_ID) %>% 
  dplyr::count() %>% ungroup() %>%
  filter(n==1)
ind = ind %>% filter(Taxonomy_ID %in% ind_pos$Taxonomy_ID) %>%
  filter(Ind==1) %>% dplyr::select(-Ind)
# Remove prefix from groups
srv = Vectorize(str_remove)
ind$group = srv(ind$group, 's.')
# Number of Indicators in each group
table(ind$group)
ind = ind %>% arrange(group)

write.csv(ind, 'Results Files/Virome/Indicator Species Statistics.csv',row.names=F)
```

```{r Plot taxa of interest}
psg = ps %>% tax_glom('Species') %>% microbiome::transform('compositional') %>% 
  psmelt() %>% 
  filter(Species %in% ind$Taxonomy_ID) %>% 
  left_join(ind %>% select(Taxonomy_ID,group) %>% unique %>% 
              rename(Species = Taxonomy_ID, Indicator = group)) %>% 
  filter(Day =='Day 28') %>% 
  left_join(maas %>% filter(Level=='Species') %>% select(Species,coef,qval)) %>% 
  mutate(symb = ifelse(qval>0.05,'',ifelse(qval>0.01,"* ",
                      ifelse(qval>0.001,'**',ifelse(qval>0.0001,'***','****'))))) %>% 
  mutate(Species = str_wrap(Species,25))

# LM-ASSOCIATED ~~~~~~~~~~~~~~~~~~~~~

# Set order for indicators
order = psg %>% filter(Diet=='LM',Indicator=='LM Day 28') %>% 
  group_by(Species) %>% summarize(Abundance = mean(Abundance)) %>% ungroup() %>% 
  arrange(-Abundance)

# Set x and y coords for the p values
max_per_path = psg %>% group_by(Species) %>% summarize(M = 2*max(Abundance)) %>% ungroup %>% 
  mutate(Species = factor(Species, levels = order$Species)) %>% 
  filter(!is.na(Species))
max_pos = max_per_path %>% 
  arrange(Species) %>% mutate(N = c(1:nrow(.))+0.25) %>% 
  mutate(M_same = max(.$M))

p1 = psg %>% filter(Indicator == 'LM Day 28') %>% droplevels() %>% 
  mutate(Species = factor(Species,levels = order$Species)) %>% 
  left_join(max_pos) %>% 
  mutate(Abundance = Abundance + min(.$Abundance[.$Abundance !=0])/2) %>% 
  ggplot(aes(Species, Abundance, fill = Group)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position=position_jitterdodge(jitter.height=0,jitter.width = 0.2)) +
  theme_classic(base_size = 16) +
  ylab('Relative Abundance') +
  xlab(NULL) +
  geom_text(aes(x=N+0.3,y=M_same,label=symb), size = 7,lineheight=0.8,hjust=2) +
  scale_y_log10(expand = expansion(mult = c(0.1, 0.1))) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

# CON-ASSOCIATED ~~~~~~~~~~~~~~~~~~~~~

# Set order for indicators
order = psg %>% filter(Diet=='CON',Indicator=='CON Day 28') %>% 
  group_by(Species) %>% summarize(Abundance = mean(Abundance)) %>% ungroup() %>% 
  arrange(-Abundance)

# Set x and y coords for the p values
max_per_path = psg %>% group_by(Species) %>% summarize(M = 2*max(Abundance)) %>% ungroup %>% 
  mutate(Species = factor(Species, levels = order$Species)) %>% 
  filter(!is.na(Species))
max_pos = max_per_path %>% 
  arrange(Species) %>% mutate(N = c(1:nrow(.))+0.25) %>% 
  mutate(M_same = max(.$M))

p2 = psg %>% filter(Indicator == 'CON Day 28') %>% droplevels() %>% 
  mutate(Species = factor(Species,levels = order$Species)) %>% 
  left_join(max_pos) %>% 
  mutate(Abundance = Abundance + min(.$Abundance[.$Abundance !=0])/2) %>% 
  ggplot(aes(Species, Abundance, fill = Group)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position=position_jitterdodge(jitter.height=0,jitter.width = 0.2)) +
  theme_classic(base_size = 16) +
  ylab('Relative Abundance') +
  xlab(NULL) +
  geom_text(aes(x=N,y=M_same,label=symb), size = 7,lineheight=0.8,hjust=1) +
  scale_y_log10(expand = expansion(mult = c(0.1, 0.1))) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

p = ggarrange(plotlist = list(p2,p1),common.legend = T,legend = 'right', widths = c(2,5))

ggsave('Plots/Virome/Fig. 3f - Viral Indicator Species.png',height = 6, width = 10.5)
```
